---
title: "Module12Guided"
output: html_document
---

```{r}
library(tidyverse)
library(faraway)
library(ROCR)
df<- wcgs

set.seed(6021)
```


```{r}
sample.data<-sample.int(nrow(df), floor(0.5*nrow(df)), replace = F)
train<-df[sample.data,]
test<-df[-sample.data,]

model<- glm(chd~age+sdp+cigs+dibep, family=binomial, data=train)

summary(model)
```
This model suggests that all behaviors outside of dibebB will increase one's log odds of having coronary heart disease with each unit increase when controlling for other variables. with dibep B only decreasing log odds by 0.5318*1 when behavior B is displayed.

### 2
```{r}
preds<- predict(model, newdata=test, type="response")

```

```{r}
rates<-ROCR::prediction(preds, test$chd)

roc_result<- ROCR::performance(rates,measure="tpr", x.measure="fpr")

plot(roc_result, main = "ROC Curve for Log Model")
lines(x=c(0,1), y = c(0,1), col="red")
```

This curve tells us that other than on the extreme ends of the model, the model performs better than randomly guessing by being above the 50-50 line for essentially the whole plot

### 3
```{r}
auc<- ROCR::performance(rates, measure = "auc")
auc@y.values
```

The AUC value agrees with my own visual conclusion that the ROC curve performs better than random guessing

### 4

```{r}
table(test$chd, preds>0.5)
```

This would suggest that I need to lower my threshold as I am not getting in True negatives nor am I getting any true positives from this output, instead I am getting a 2x1 matrix where every outcome is False.

### 5
I think in this case I would disagree, the Matrix is giving an output where the FPR and FNR = 1 meaning that all outcomes from the model at this threshold are wrong. making it perform worse than random guessing.

### 6 
The threshold should be lowered. This will allow for the model to predict that patients have chd. With coronary heart disease the model should have a certain degree of sensitivity due to the risk it puts on a life that using these measurements we would want their risk factor to be caught and allowing for them to then be checked by a professional. However, to a degree you will want to cap this sensitivity to streamline patient care if this were a screening survey for potential risk populations and not make them end up having to test every patient they see. 

### 7
```{r}
summary(preds)
```
```{r}
table(test$chd, preds>0.04312)
```
accuracy-> 75.02%
FPR-> 0.562%
TPR-> 23.92% *clearly can improve* probably closer to minimum threshold

### 8
I think what is happening is like with much of medical screening this test needs to be HIGHLY sensitive
