---
title: "Guided Set 2"
author: "Michael Puchalski"
date: "2025-01-28"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(tidyverse)
df<-read.csv("new_students.csv", header = TRUE)
```


#Question 1
Frequency Table of the number of students in each level of GPA.cat then arrange if needed

```{r}
df$GPA.cat<- factor(df$GPA.cat, levels=c("low","moderate","high"))
# levels(df$GPA.cat) checks the order of levels
table(df$GPA.cat)
```


#Question 2
Create a bar chart for this data

```{r}
df_clean<-df%>%
  drop_na(GPA.cat)
ggplot(df_clean, aes(x=GPA.cat))+
  geom_bar() +
  theme(axis.text.x = element_text(angle=0),
        plot.title = element_text(hjust=0.5))+
  labs(x="GPA Category", y = "Student Count", title = "GPA distribution")
```

#Question 3
Create a bar chart with proportions
```{r}
prop_df<-df_clean%>%
  group_by(GPA.cat)%>%
  summarize(Counts=n())%>%
  mutate(Percent = Counts/nrow(df_clean))
prop_df
ggplot(prop_df, aes(x=GPA.cat, y=Percent))+
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 0),
        plot.title = element_text(hjust = 0.5))+
  labs(x="GPA Category", y = "Student Proportion", title = "GPA distribution proportions")

```

#Question 4
Two Way table for the number of male and female students and the GPA category

```{r}
two_way_table<-table(df_clean$Gender, df_clean$GPA.cat)
two_way_table
```

#Question 5
Produce a percentage table for the proportion of GPA categories
```{r}
round(prop.table(two_way_table, 1)*100, 2)
round(prop.table(two_way_table, 2)*100, 2)
```
According to these tables it shows that females make up a larger percentage of the moderate and high GPA.categories with men making up just over half of the low GPA category. Along with this, men show to have a higher grouping in the low vs high end where the opposite is true for females.

#Question 6
make a bivariute bar chart
```{r}
ggplot(df_clean, aes(x = GPA.cat, fill = Gender))+
  geom_bar(position = "fill")+
  labs(x="GPA Category", y = "Proportion", title = "GPA by category and gender")
```

#Question 7 
Split this further by smoker vs nonsmoker
```{r}
ggplot(df_clean, aes(x = GPA.cat, fill = Smoke))+
  geom_bar(position = "fill")+
  facet_wrap(~Gender)
```

This chart would probably be better as a count chart, but the way to give the most information is by allowing it to wrap by gender which shows that the more successful students both male and female have a lower rate of being smokers than in the lower GPA categories.That being said, it is a chart that needs context to allow it to be a proper visualization.

#Question 8
Creat GPA vs study hours scatter plot
```{r}
ggplot(df_clean, aes(x = StudyHrs, y = GPA))+
  geom_point()
```
This chart shows that there is no guarantee of success, but that when more than 30 hours are spent studying, the lowest GPA is a 2.75. So, while increasing study hours does not guarantee an increase in GPA, studying more can correlate. Along with this, it shows that su=ome students can succeed with relatively low study hrs per week.

#Question 9 
Make the scatter plot include party days too
```{r}
ggplot(df_clean, aes(x = PartyNum, y = StudyHrs, color = GPA))+
  geom_point()
```

#Question 10
same thing plus smoking

```{r}
ggplot(df_clean, aes(x = PartyNum, y = StudyHrs, color = GPA, size = Smoke))+
  geom_point()
```

