---
title: "Guided Set Mod 9"
author: "Michael Puchalski"
date: "2025-03-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(leaps)
Data<- read.table("nfl.txt", header=T)
```

### 1.

```{r}
allreg<-regsubsets(y ~., data=Data, nbest=2)
summary(allreg)
```
Based on R2, best model for 1-predictors is that with x8 as the 1-predictor with the second best being that with x1.
```{r}
names(summary(allreg))
```


```{r}
which.max(summary(allreg)$adjr2)
```
```{r}
which.min(summary(allreg)$cp)
```

```{r}
which.min(summary(allreg)$bic)
```

### 2 forward elimination
```{r}
#intercept only model
regnull<-lm(y~1, data=Data)
# all predictors
regfull<-lm(y~., data=Data)
```

```{r}
# forward selection
step(regnull, scope=list(lower=regnull,upper=regfull), direction ="forward")
```

### 3 backward elimination

```{r}
step(regfull, scope=list(lower=regnull,upper=regfull),direction="backward")
```

### 4 stepwise regression
```{r}
step(regfull, scope=list(lower=regnull,upper=regfull),direction="both")

```
### 5 PRESS computation function
```{r}
compute_PRESS<- function(model){
  h<- lm.influence(model)$hat
  e<-residuals(model)
  PRESS<- sum((e/(1-h))^2)
  return(PRESS)
}
```


### 6 applying PRESS
```{r}
model<-lm(y~x2+x7+x8, data = Data)
PRESS_val<- compute_PRESS(model)
PRESS_val
```


```{r}
R2<- summary(model)$r.squared
R2
```
```{r}
summary(model)
anova(model)
```

```{r}
SST<-326.964
R2_Pred<- 1- PRESS_val/SST
R2_Pred
R2
```
Values are close suggesting the model is a decent generalization of the model where we will likely not suffer from overfitting because it is smaller the R2 as expected, but not greatly so.

trying to Commit
