---
title: "HW01"
author: "Michael Puchalski"
date: "2025-01-25"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(dplyr)
df<-read.csv("UScovid.csv")
dim(df)
colnames(df)
class(df$date)
str(date)
typeof(df$date)
```
##Question 1a (cleaning up the data)
We are interested in the data at the most recent date, June 3 2021. Create a data
 frame called latest that:
- has only rows pertaining to data from June 3 2021,
- removes rows pertaining to counties that are “Unknown”,
- removes the column date and fips,
- is ordered by county and then state alphabetically
Display the first 6 rows of the data using head().
```{r}
latest<-df%>%
  mutate(date = as.Date(date)) %>%
  filter(date == as.Date('2021-06-03'))%>% # Filters the data frame to only include the selected date. Doing as.Date a second time may be redundant
  filter(!is.na(county) & county!= "Unknown")%>% # This removes rows with the null value from the specific column
  select(-date, -fips)%>% # Using '-' sign allows us to remove the columns that are selected
  arrange(county,state) # This allows us to sort alphabetically
head(latest, n = 6)
```
## Question 1b
Calculate the case fatality rate (number of deaths divided by number of cases, and call it death.rate) for each county. Report the case fatality rate as a percent and round to two decimal places. Add death.rate as a new column to the data frame latest. Display the first 6 rows of the data frame latest.
```{r}
latest<-latest%>%
  mutate(death.rate = round((deaths/cases) * 100, 2))
# round function allows us to define after the comma what decimal we want to round to
head(latest, n = 6)
```

## Question 1c
Display the counties with the 10 largest number of cases. Be sure to also display the number of deaths and case fatality rates in these counties, as well as the state the counties belong to.
```{r}
top_ten_cases<-latest %>%
  arrange(desc(cases))
head(top_ten_cases, n = 10)
```

##Question 1d
Display the counties with the 10 largest number of deaths. Be sure to also display the number of cases and case fatality rates in these counties, as well as the state the counties belong to
```{r}
top_ten_deaths<-latest%>%
  arrange(desc(deaths))
head(top_ten_deaths, n = 10)

```

##Question 1e
Display the counties with the 10 highest case fatality rates. Be sure to also display the number of cases and deaths in these counties, as well as the state the counties belong to. Is there something you notice about these counties?
```{r}
top_ten_death_rate<-latest %>%
  arrange(desc(death.rate))
head(top_ten_death_rate, n = 10)
```
What is noticeable about the counties is that they all seem to be rural counties where the 
overall incedence of infection is lower, but that also means that that an instance of death is going to have a greater impact.

##Question 1f
 Display the counties with the 10 highest case fatality rates among counties with at least 100,000 cases. Be sure to also display the number of cases and deaths in these counties, as well as the state the counties belong to.

```{r}
larger_set_top_ten<-latest%>%
  filter(cases>= 100000)%>%
  arrange(desc(death.rate))
head(larger_set_top_ten, n = 10)
```

##Question 1g
Display the number of cases, deaths, case fatality rates for the following counties:
 i. Albemarle, Virginia
 ii. Charlottesville city, Virginia
```{r}
local_rate<-latest %>%
  filter(state == "Virginia")%>%
  filter(county == "Albemarle" | county == "Charlottesville city")
local_rate
```

#Question 2
For this question, we focus on data at the state level. Note that the dataset has data
 on the 50 states, plus DC, Puerto Rico, Guam, Northern Mariana Islands, and the
 Virgin Islands. For the purpose of this question, we will consider DC, Puerto Rico,
 Guam, Northern Mariana Islands, and the Virgin Islands, as “states” as well.
##Question 2a
We are interested in the data at the most recent date, June 3 2021. Create a data
 frame called state.level that:
- has 55 rows: 1 for each state, DC, and territory
- has 3 columns: name of the state, number of cases, number of deaths
- is ordered alphabetically by name of the state
Display the first 6 rows of the data frame state.level.
```{r}
state.level<-df%>%
  group_by(state)%>%
  summarize(
    cases = sum(cases, na.rm = TRUE),
    deaths = sum(deaths, na.rm = TRUE) # inserted in na.rm = TRUE because without it, Puerto Rico has no death count im the summary statistic
  ) %>%
  arrange(state)
head(state.level, n = 6)
```
##Question 2b
Calculate and add the state.rate case fatality rate
```{r}
state.level<-state.level%>%
  mutate(state.rate = round((deaths/cases) * 100, 2))
head(state.level, n = 6)
```

##Question 2c and 2d
What is the fatility rate in Virginia and Puerto Rico?
```{r}
filter(state.level, state == "Virginia" | state == "Puerto Rico")
```
According to my output, the death rate for Virginia is 1.67 and 1.43 for Puerto Rico. This was only attainable after going back and making the summary statistics ignore NA values during summation. 

##Question 2e
Which states have the 10 highest case fatality rates?
```{r}
top_fatality<-state.level%>%
  arrange(desc(state.rate))
head(top_fatality, n = 10)
```
In order: New York, New Jersey, Conneticut, DC, Pennsylvania, Michigan, Louisiana, Mississippi, Maryland
##Question 2f
Which states have the 10 lowest fatality rates?
```{r}
low_fatality<-state.level%>%
  arrange(state.rate)
head(low_fatality, n = 10)
```
In order: Alaska, Utah, Nebraska, Idaho, Wyoming, Wisconsin, Virgin Islands, Oklahoma, Montana, Kentucky

##Question 2g
Export this dataset as a .csv file named stateCovid.csv. We will be using this file for the next homework.
```{r}
write.csv(state.level, "stateCovid.csv", row.names = FALSE)
```

