---
title: "Project 1 Markdown"
author: "Sean Hersee, Gregory Miller, Sree Prabhav Bandakavi, Michael Puchalski"
date: "2025-03-23"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(MASS)
library(scales)
library(ggsignif)
library(knitr)
library(kableExtra)
data<-read.csv('diamonds4.csv')
options(scipen = 999)
```

## Price against Carat Linear Regression

##### Step 1: Simple Linear Regression & Plot

```{r}
slr<-lm(data$price~data$carat,data=data)
ggplot2::ggplot(data, aes(x=carat, y=price))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="Carat Weight",y="Price",title="Scatterplot of Sales Price against Carat Weight")+
  scale_y_continuous(labels = dollar_format())
par(mfrow = c(2, 2))
plot(slr)
```

Assumption 1 is not met: Points are not evenly scattered on either side of the line on the scatterplot\
Assumption 2 is not met: The average value of residuals (red) line on the Residuals Plot is not along the horizontal\
axis 0 and the vertical spread of the data points varies as we move left to right on the chart

##### Step 2: Determine Transformation of the response variable

```{r}
MASS::boxcox(slr, lambda = seq(-.1, .4, 1/10))
```

Although Lambda of 0 is not within the bounds of the confidence interval, I will choose a Lambda of 0 as that will allow for a log transformation which will lead to comprehensible results

##### Step #3: Apply Log to Response Variable and view results

```{r}
ystar<-log(data$price)
data<-data.frame(data,ystar)
slr.trans1<-lm(ystar~carat,data=data)
ggplot2::ggplot(data, aes(x=carat, y=ystar))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="Carat Weight",y="log(Price)",title="Scatterplot of Sales Price against Carat Weight")
par(mfrow = c(2, 2))
plot(slr.trans1)
```

These results appear logarithmic annd indicate that the predictor variable also needs to be transformed using a log transformation

##### Step #4: Apply Log to the Predictor and view results

```{r}
xstar<-log(data$carat)
data<-data.frame(data,xstar)
slr.trans2<-lm(ystar~xstar,data=data)
ggplot2::ggplot(data, aes(x=xstar, y=ystar))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="log(Carat Weight)",y="log(Price)",title="Scatterplot of Sales Price against Carat Weight")
par(mfrow = c(2, 2))
plot(slr.trans2)
summary(slr.trans2)
```

This has resulted in a useful data set where all assumptions of linear regressions are met\
$y^*=8.521 + 1.944x^*$

## Effect of price on carat, clarity, color, and cut

```{r}

alpha_values3 <- c("J" = 0.3, "I" = 0.4, "H" = 0.5, "G" = 0.6, "F" = 0.7, "E" = 0.8, "D" = 1.0)
data %>%
  group_by(color) %>%
  summarise(avg_price = mean(price, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(color, levels = names(alpha_values3)), 
             y = avg_price, 
             alpha = factor(color, levels = names(alpha_values3)))) +  
  geom_col(fill = "dodgerblue4") +  
  scale_y_continuous(labels = dollar_format()) +
  scale_alpha_manual(values = alpha_values3) + 
  labs(title = "Average Diamond Price by Color",
       x = "Color",
       y = "Average Price",
       alpha = "Color") + 
  theme_minimal()

alpha_values <- c("SI2" = .2, "SI1" = 0.4, "VS2" = 0.5, "VS1" = 0.6, "VVS2" = 0.7, "VVS1" = 0.8, "IF" = 0.9, "FL" = 1)
data %>%
  group_by(clarity) %>%
  summarise(avg_price = mean(price, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(clarity, levels = names(alpha_values)), 
             y = avg_price, alpha = factor(clarity, levels = names(alpha_values)))) +
  geom_col(fill = "dodgerblue4") +
  scale_alpha_manual(values = alpha_values) +  
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Average Diamond Price by Clarity",
       x = "Clarity",
       y = "Average Price",
       alpha = "Clarity") +  # Proper legend title
  theme_minimal()

alpha_values2 <- c("Good" = 0.2, "Very Good" = 0.6, "Ideal" = 0.8, "Astor Ideal" = 1.0)
data %>%
  group_by(cut) %>%
  summarise(avg_price = mean(price, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(cut, levels = names(alpha_values2)),
             y = avg_price, 
             alpha = factor(cut, levels = names(alpha_values2)))) + 
  geom_col(fill = "dodgerblue4") +
  scale_alpha_manual(values = alpha_values2) +
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Average Diamond Price by Cut",
       x = "Cut",
       y = "Average Price",
       alpha = "Cut") +  # Proper legend title
  theme_minimal()


data <- data %>%
  mutate(carat_group = case_when(
    carat < 1 ~ "< 1.00 carat",
    carat >= 1 & carat < 2 ~ "1.00 - 1.99 carats",
    carat >= 2 & carat < 3 ~ "2.00 - 2.99 carats",
    carat >= 3 & carat < 4 ~ "3.00 - 3.99 carats",
    carat >= 4 & carat < 5 ~ "4.00 - 4.99 carats",
    carat >= 5 ~ "5.00+ carats"
  ))

alpha_values4 <- c("< 1.00 carat" = 0.3, 
                  "1.00 - 1.99 carats" = 0.4, 
                  "2.00 - 2.99 carats" = 0.5, 
                  "3.00 - 3.99 carats" = 0.6, 
                  "4.00 - 4.99 carats" = 0.7, 
                  "5.00+ carats" = 0.8)

data %>%
  group_by(carat_group) %>%
  summarise(avg_price = mean(price, na.rm = TRUE)) %>%
  ggplot(aes(x = carat_group, y = avg_price, alpha = carat_group)) +
  geom_col(fill = "dodgerblue4") +  # Keep all bars the same color
  scale_y_continuous(labels = dollar_format()) +
  scale_alpha_manual(values = alpha_values4) +  
  labs(title = "Average Diamond Price by Carat",
       x = "Carat Group",
       y = "Average Price",
       alpha = "Carat Group") +  # Proper legend title
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```

##Buy shy to save money - Diamond Price by Carat/Diamond Price by Carat and Color Group

```{r}
ggplot(data, aes(x = carat, y = price)) +
  geom_line(stat = "summary", fun = mean, size = 1) + 
  labs(title = "Average Diamond Price by Carat",
       x = "Carat",
       y = "Average Price ($)") +  
  scale_y_continuous(labels = dollar_format()) + 
  theme_minimal()

data <- data %>%
  mutate(color_group = case_when(
    color %in% c('D', 'E', 'F') ~ "Colorless",
    color %in% c('G', 'H', 'I', 'J') ~ "Near-colorless",
    color %in% c('K') ~ "Faint color",
    TRUE ~ as.character(color)
  ))

ggplot(data, aes(x = carat, y = price, color = color_group)) +
  geom_line(stat = "summary", fun = mean, size = 1) + 
  labs(title = "Average Diamond Price by Carat and Color Group",
       x = "Carat",
       y = "Average Price ($)",
       color = "Color Group") +  
  scale_y_continuous(labels = dollar_format()) +  
  theme_minimal()

```

## Analyzing the Impact of Cut across Carat Group

```{r}
data.sub<-data %>% 
  filter(carat<1.4)

data.sub<-data.sub %>% 
  mutate(carat.group=cut(x=data.sub$carat,breaks=c(0.2,.6,1,1.4),labels=c("0.2 - 0.6 Carats","0.6 - 1.0 Carats","1.0 - 1.4 Carats")))
data.sub$cut<-factor(data.sub$cut,levels=c("Good","Very Good","Ideal","Astor Ideal"))

ggplot(data.sub,aes(x=cut,y=ystar))+
  geom_boxplot(aes(fill=carat.group,alpha=cut))+
  facet_wrap(~carat.group)+
  labs(title="Impact of Cut on Price by Carat Group",y='log(Price)',x="Cut")+
   scale_alpha_discrete(
     range = c(0.3, 0.9),
     guide = guide_legend(override.aes = list(fill = "black")))+
  scale_fill_manual(values = c("goldenrod2", "forestgreen", "dodgerblue4"))
```

The summary for Color

```{r}
model <- lm(formula = price ~ color, data = data)
summary(model)
model2 <- lm(formula = price ~ cut, data = diamonds)
summary(model2)
```

Table format:

|             | Estimate | Std. Error | t value | Pr(\>\|t\|) |        |
|-------------|----------|------------|---------|-------------|--------|
| (Intercept) | 10525.2  | 1672.2     | 6.294   | 4.32e-10    | \*\*\* |
| colorE      | -619.3   | 2448.3     | -0.253  | 0.8004      |        |
| colorF      | -4320.6  | 2322.1     | -1.861  | 0.0630      | .      |
| colorG      | -5953.5  | 2391.6     | -2.489  | 0.0129      | \*     |
| colorH      | -2726.6  | 2589.9     | -1.053  | 0.2926      |        |
| colorI      | -5745.8  | 2502.5     | -2.296  | 0.0218      | \*     |
| colorJ      | -6591.1  | 3037.8     | -2.170  | 0.0302      | \*     |

|              | Estimate | Std. Error | t value | Pr(\>\|t\|) |
|--------------|----------|------------|---------|-------------|
| (Intercept)  | 5851.6   | 5397.3     | 1.084   | 0.279       |
| cutGood      | 3615.7   | 6091.9     | 0.594   | 0.553       |
| cutIdeal     | 637.4    | 5469.8     | 0.117   | 0.907       |
| cutVery Good | 1906.1   | 5536.8     | 0.344   | 0.731       |

```{r}
ggplot(data, aes(x = color, y = price, fill = color)) +
  geom_boxplot(outlier.shape = NA) +
  coord_trans(y = "log10") +
  geom_signif(comparisons = list(c("D", "G")), y_position = 10000, annotation = "*", tip_length = 0.01) +
  geom_signif(comparisons = list(c("D", "I")), y_position = 15000, annotation = "*", tip_length = 0.01) +
  geom_signif(comparisons = list(c("D", "J")), y_position = 20000, annotation = "*", tip_length = 0.01) +
  labs(title = "Diamond Prices by Color (Log Scale with Significance)",
       y = "Price (log scale)",
       x = "Color") +
  theme_minimal()
```

```{r}
data<-data %>% # used to make binning of data more uniform
mutate(clarity_group = case_when(
    clarity %in% c('SI1', 'SI2')~"SI",
    clarity %in% c('VS1', 'VS2')~"VSI",
    clarity %in% c('VVS1', 'VVS2')~"VVSI",
    clarity %in% c('IF')~"IF",
    clarity %in% c('FL')~"FL",
    TRUE ~ as.character(clarity)
  ))
data<- data %>% # pulled from code format from Greg's data manipulation
  mutate(carat_group = case_when(
    carat < 1 ~ "less than 1.00",
    carat >= 1 & carat < 2 ~ "1.00 - 1.99",
    carat >= 2 & carat < 3 ~ "2.00 - 2.99",
    carat >= 3 & carat < 4 ~ "3.00 - 3.99",
    carat >= 4 & carat < 5 ~ "4.00 - 4.99",
    carat >= 5 ~ "5.00+"
  ))
```

### Clarity compared with Cut and Carat

```{r}
data<-data %>% # used to make binning of data more uniform
  mutate(clarity_group = case_when(
    clarity %in% c('SI1', 'SI2')~"SI",
    clarity %in% c('VS1', 'VS2')~"VSI",
    clarity %in% c('VVS1', 'VVS2')~"VVSI",
    clarity %in% c('IF')~"IF",
    clarity %in% c('FL')~"FL",
    TRUE ~ as.character(clarity)
  ))

```

```{r}
ggplot(data, aes(x = factor(clarity_group, levels=c("FL","IF", "VVSI", "VSI", "SI")), fill = factor(cut, levels = c("Astor Ideal", "Ideal", "Very Good", "Good")))) +
  geom_bar(alpha = 0.7, position = "stack") +
  facet_wrap(~ carat_group, scales = "free_y")+
  labs(title = "Clarity Groupings Compared with Cut and Carat", 
       x = "Clarity Group",
       y = "Count",
       fill = "Cut Quality") +
  theme_bw() +  # theme for better readability
  theme(
      plot.title   = element_text(hjust= 0, size = 12),
      axis.text.x  = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8),
      axis.text.y  = element_text(size = 12)
    )
# A table summarizing counts for clarity_group, carat_group, and cut
summary_table <- as.data.frame(
  table(data$clarity_group, data$carat_group, data$cut)
)

# Renamimg the columns of the table
colnames(summary_table) <- c("Clarity Group", "Carat Group", "Cut", "Count")

```

### Tables to correspond to the facet wrapped

```{r}
# Filtered and grouped by carat_group, then clarity_group
filtered_data <- data %>%
  group_by(carat_group, clarity_group, cut) %>%
  summarise(Count = n(), .groups = "drop") %>%
  filter(Count != 0)

# Data by carat_group
facet_tables <- split(filtered_data, filtered_data$carat_group)

#summary tables for each carat_group
summary_tables <- lapply(facet_tables, function(group_data) {
  group_data <- group_data %>%
    group_by(clarity_group, cut) %>%
    summarise(Count = sum(Count), .groups = "drop") %>%
    mutate(Percentage = round((Count / sum(Count) * 100), 2)) %>%
    rename(
      Clarity = clarity_group,  
      Cut = cut
    )
  return(group_data)
})
# exportable tables

for (carat in names(summary_tables)) {
  output_file <- paste0("table_", carat, ".html")
  cat("\n### Carat Group:", carat, "\n")
  
  summary_tables[[carat]] %>%
    kable(
      caption = paste("Table for Carat Group:", carat),
      format = "html"
    ) %>%
    kable_styling(
      bootstrap_options = c("striped", "hover", "condensed"), 
      full_width = FALSE,
      position = "center" 
    ) %>%
    add_header_above(c(" " = 2, "Summary" = 2))%>%
    save_kable(file = output_file)
}
```


