---
title: "ProjectPlots"
output: html_document
---

```{r}
library(tidyverse)
library(MASS)
library(scales)
data<-read.csv('diamonds4.csv')
```

### Simple linear regression: Price against Carat

Pulled from Sean's work for consistency. Plots consistent with the work from a log base 0.3 modification. Will follow the progress of Sean's regression model.

```{r}
slr<-lm(data$price~data$carat,data=data)
ggplot2::ggplot(data, aes(x=carat, y=price))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="Carat Weight",y="Price",title="Scatterplot of Sales Price against Carat Weight")
par(mfrow = c(2, 2))
plot(slr)
```

Assumption 1 is not met: Points are not evenly scattered on either side of the line on the scatterplot\
Assumption 2 is not met: The average value of residuals (red) line on the Residuals Plot is not along the horizontal\
axis 0 and the vertical spread of the data points varies as we move left to right on the chart

##### Transformation of response variable

```{r}
MASS::boxcox(slr, lambda = seq(-.1, .4, 1/10))
```

Utilizing a natural log as it is common practice to round down for more effective interpretation if it is possible to round down to 0 in the confidence interval values.

```{r}
ystar<-log(data$price)
data<-data.frame(data,ystar)
slr.trans1<-lm(ystar~carat,data=data)
ggplot2::ggplot(data, aes(x=carat, y=ystar))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="Carat Weight",y="Price",title="Scatterplot of Sales Price against Carat Weight")
par(mfrow = c(2, 2))
plot(slr.trans1)
```

Predictor variable now needs to be normalized via a logarithmic transformation as well to meet assumption 1.

```{r}
xstar<-log(data$carat)
data<-data.frame(data,xstar)
slr.trans2<-lm(ystar~xstar,data=data)
ggplot2::ggplot(data, aes(x=xstar, y=ystar))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="Carat Weight",y="Price",title="Scatterplot of Sales Price against Carat Weight")
par(mfrow = c(2, 2))
plot(slr.trans2)
```

```{r}
data<-data %>% # used to make binning of data more uniform, from Greg
  mutate(clarity_group = case_when(
    clarity %in% c('SI1', 'SI2')~"Slightly Included Diamonds",
    clarity %in% c('VS1', 'VS2')~"Very Slightly Included Diamonds",
    clarity %in% c('VVS1', 'VVS2')~"Very, Very Slightly Included Diamonds",
    clarity %in% c('IF')~"Internally Flawless Diamonds",
    clarity %in% c('FL')~"Flawless Diamonds",
    TRUE ~ as.character(clarity)
  ))
data<- data %>% # pulled from code format from Greg's data manipulation
  mutate(carat_group = case_when(
    carat < 1 ~ "< 1.00",
    carat >= 1 & carat < 2 ~ "1.00 - 1.99",
    carat >= 2 & carat < 3 ~ "2.00 - 2.99",
    carat >= 3 & carat < 4 ~ "3.00 - 3.99",
    carat >= 4 & carat < 5 ~ "4.00 - 4.99",
    carat >= 5 ~ "5.00+"
  ))
```

```{r}
# Plot formatting incorporating personal work with formatting suggestions from Sean and Greg
data %>%
  group_by(clarity) %>%
  summarise(avg_price = mean(price)) %>%
  ggplot(aes(x = factor(clarity, levels = c("FL", "IF", "VVS1", "VVS2", "VS1", "VS2", "SI1", "SI2", "I1")), y = avg_price, fill = clarity)) +
  scale_y_continuous(labels = dollar_format()) +  # Format as $
  geom_col() +
  geom_text(aes(label = dollar_format()(avg_price)), 
          vjust = -0.3,  # Position text slightly above bars
          size = 3) +    # Adjust text size
  labs(title = "Average Diamond Price by Clarity",
       x = "Clarity",
       y = "Average Price") +
  theme_minimal()

ggplot(data, aes(x = clarity, y = price, fill = clarity)) +
  geom_boxplot() +
  facet_wrap(~ carat_group, scales = "free_y") +  # Allow different y-scales
  scale_y_continuous(labels = dollar_format()) +  # Format as $
  labs(title = "Diamond Price by Clarity and Carat Group",
       x = "Clarity",
       y = "Price") +
  theme_minimal()

ggplot(data, aes(x = carat, y = ystar, color = clarity_group)) +
  geom_line(stat = "summary", fun = mean, linewidth = 1) + 
  labs(title = "Average Diamond Price by Carat and Clarity Group",
       x = "Carat",
       y = "Price ($)",
       color = "Clarity Group") +  # Legend label
  scale_y_continuous(labels = dollar_format()) +  # Format as $
  theme_minimal()
```

### Cut to Clarity

```{r}
data<-data %>% 
  mutate(carat.group=cut(x=carat,breaks= unique(c(quantile(data$carat,probs = seq(0, 1, by =0.25),include.lowest=TRUE)))))%>%
  mutate(carat.group = coalesce(as.character(carat.group), "Other"))
data$cut<-factor(data$cut,levels=c("Good","Very Good","Ideal","Astor Ideal"))

# ggplot(data,aes(x=clarity_group))+
 # geom_histogram()+
  # xlim(0,2.5)

ggplot(data,aes(x=xstar,y=price))+
  geom_violin()+
  facet_wrap(~clarity_group)


ggplot(data,aes(x=xstar,y=av,fill=clarity_group))+
  geom_point(shape=21,alpha=.7)
```

```{r}
alpha_values <- c("SI2" = .2, "SI1" = 0.4, "VS2" = 0.5, "VS1" = 0.6, "VVS2" = 0.7, "VVS1" = 0.8, "IF" = 0.9, "FL" = 1)
```

```{r}
ggplot(data,aes(x=carat.group, fill=color))+
  geom_bar(alpha = 0.7, position = "stack") + # Default stacked bar behavior
  geom_text(stat = "count", aes(label = ..count..), 
            position = position_stack(vjust = 0.5), # Center labels within blocks
            size = 3)+
  facet_wrap(~clarity_group)

data %>%
  group_by(clarity) %>%
  summarise(avg_price = mean(price, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(clarity, levels = names(alpha_values)), 
             y = avg_price, alpha = factor(clarity, levels = names(alpha_values)))) +
  geom_col(fill = "dodgerblue4") +
  scale_alpha_manual(values = alpha_values) +  
  scale_y_continuous(labels = dollar_format()) +
  labs(title = "Average Diamond Price by Clarity",
       x = "Clarity",
       y = "Average Price",
       alpha = "Clarity") +  # Proper legend title
  theme_minimal()  
```

```{r}
# Get unique facets (clarity_group levels) 
facets <- unique(data$clarity_group)

# Create a plot for each facet
for (facet in facets) {
  # Filter data for the current facet
  facet_data <- data %>% filter(clarity_group == facet)
  
  # Creating the plots
  p <- ggplot(facet_data, aes(x = carat_group, fill = cut)) +
    geom_bar(alpha = 0.7, position = "stack") +
    geom_text(stat = "count", aes(label = after_stat(count)), 
              position = position_stack(vjust = 0.5),
              size = 3) +
    labs(title = paste("Clarity Group:", facet), x = "Carat Group", y = "Count")
  print(p)
}
```

```{r}
for (facet in facets) {
  # Filter data for the current facet
  facet_data <- data %>% filter(clarity_group == facet)
  
  # Create the plot
  q <- ggplot(facet_data, aes(x = carat.group, fill = color)) +
    geom_bar(alpha = 0.7, position = "stack") +
    geom_text(stat = "count", aes(label = ..count..), 
              position = position_stack(vjust = 0.5),
              size = 4,          # Increase text size
              color = "white",   # Use white text for contrast (adjust if needed)
              fontface = "bold") +  # Make text bold
    labs(title = paste("Clarity Group:", facet), 
         x = "Carat Group", 
         y = "Count") +
    theme_bw() +  # A clean theme for better readability
    theme(
      plot.title   = element_text(size = 16, face = "bold"),
      axis.text.x  = element_text(size = 12),
      axis.text.y  = element_text(size = 12)
    )
  
  # Print the plot to display it in your plotting window
  print(q)
}
```

```{r}
print(data$carat.group)
```

```{r}
data %>%
  group_by(clarity) %>%
  summarise(avg_price = mean(price)) %>%
  ggplot(aes(x = factor(clarity, levels = c("FL", "IF", "VVS1", "VVS2", "VS1", "VS2", "SI1", "SI2", "I1")), y = avg_price, fill = clarity)) +
  scale_y_continuous(labels = dollar_format()) +  # Format as $
  geom_col(fill="dodgerblue4") +
  geom_text(aes(label = dollar_format()(avg_price)), 
          vjust = -0.3,  # Position text slightly above bars
          size = 3) +    # Adjust text size
  labs(title = "Average Diamond Price by Clarity",
       x = "Clarity",
       y = "Average Price") +
  theme_minimal()

ggplot(data, aes(x = clarity, y =, fill = carat_group)) +
  geom_boxplot() +
  facet_wrap(~ cut, scales = "free_y") +  # Allow different y-scales
  scale_y_continuous(labels = dollar_format()) +  # Format as $
  labs(title = "Diamond Price by Clarity and Carat Group",
       x = "Clarity",
       y = "Price") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 8))  # Rotate x-axis labels

ggplot(data, aes(x = carat, y = ystar, color = clarity_group)) +
  geom_line(stat = "summary", fun = mean, size = 1) + 
  labs(title = "Average Diamond Price by Carat and Clarity Group",
       x = "Carat",
       y = "Price ($)",
       color = "Clarity Group") +  # Legend label
  scale_y_continuous(labels = dollar_format()) +  # Format as $
  theme_minimal()
```

