---
title: "Project 1"
author: "Sean Hersee"
date: "2025-03-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(MASS)
data<-read.csv('diamonds4.csv')
```

# Claims:  
1) Cut is the most important factor in determining diamond quality  
  -Higher cut quality will make for much more expensive diamonds  
    Ranking:  
    Ideal: Top 3%  
    Very Good: Top 15%  
    Good : Top 25%  
2) Color is the 2nd most important characteristic  
  -colorless (higher color grading) will be more expensive  
     -scale in data set goes from D (colorless) to K (faintly colored)  
3) Clarity is an important but not most impactful factor  
  -Scale runs from slightly included to flawless  
4) Carat is the weight of the diamond, not the size  
  -Carat will have a positive linear relationship with price, but is not an important factor when considering the  
  impact of cut and shape  


## Price against Carat Linear Regression

##### Step 1: Simple Linear Regression & Plot
```{r}
slr<-lm(data$price~data$carat,data=data)
ggplot2::ggplot(data, aes(x=carat, y=price))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="Carat Weight",y="Price",title="Scatterplot of Sales Price against Carat Weight")
par(mfrow = c(2, 2))
plot(slr)
```

Assumption 1 is not met: Points are not evenly scattered on either side of the line on the scatterplot  
Assumption 2 is not met: The average value of residuals (red) line on the Residuals Plot is not along the horizontal  
axis 0 and the vertical spread of the data points varies as we move left to right on the chart

##### Step 2: Determine Transformation of the response variable
```{r}
MASS::boxcox(slr, lambda = seq(-.1, .4, 1/10))
```

Although Lambda of 0 is not within the bounds of the confidence interval, I will choose a Lambda of 0 as that will allow
for a log transformation which will lead to comprehensible results

##### Step #3: Apply Log to Response Variable and view results
```{r}
ystar<-log(data$price)
data<-data.frame(data,ystar)
slr.trans1<-lm(ystar~carat,data=data)
ggplot2::ggplot(data, aes(x=carat, y=ystar))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="Carat Weight",y="Price",title="Scatterplot of Sales Price against Carat Weight")
par(mfrow = c(2, 2))
plot(slr.trans1)
```

These results appear logarithmic annd indicate that the predictor variable also needs to be transformed using a log
transformation

##### Step #4: Apply Log to the Predictor and view results
```{r}
xstar<-log(data$carat)
data<-data.frame(data,xstar)
slr.trans2<-lm(ystar~xstar,data=data)
ggplot2::ggplot(data, aes(x=xstar, y=ystar))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="Carat Weight",y="Price",title="Scatterplot of Sales Price against Carat Weight")
par(mfrow = c(2, 2))
plot(slr.trans2)
```

This has resulted in a useful data set where all assumptions of linear regressions are met


#Cut to Price across multiple carat intervals (0.5, 1.0, 1.5, 2.0)
```{r}
data<-data %>% 
  mutate(carat.group=cut(x=data$carat,breaks=c(quantile(data$carat,probs = seq(0, 1, by =0.25),include.lowest=TRUE))))
data$cut<-factor(data$cut,levels=c("Good","Very Good","Ideal","Astor Ideal"))

#data<-data %>% 
#  mutate(carat.group=cut(x=data$carat,breaks=c(0.2,.5,1,1.5,2,Inf)))
#data$cut<-factor(data$cut,levels=c("Good","Very Good","Ideal","Astor Ideal"))

ggplot(data,aes(x=carat))+
  geom_histogram()+
  xlim(0,2.5)

ggplot(data,aes(x=cut,y=price))+
  geom_violin()+
  ylim(0,7500)

ggplot(data,aes(x=cut,y=ystar))+
  geom_violin()+
  facet_wrap(~carat.group)

ggplot(data,aes(x=xstar,y=ystar,fill=cut))+
  geom_point(shape=21,alpha=.7)

```


```{r}
data.sub<-data %>% 
  filter(carat<1.4)

data.sub<-data.sub %>% 
  mutate(carat.group=cut(x=data.sub$carat,breaks=c(0.2,.6,1,1.4)))
data.sub$cut<-factor(data.sub$cut,levels=c("Good","Very Good","Ideal","Astor Ideal"))

ggplot(data.sub,aes(x=cut,y=ystar))+
  geom_violin()+
  facet_wrap(~carat.group)

ggplot(data.sub,aes(x=carat,y=ystar,fill=cut))+
  geom_point(shape=23,alpha=.6)

ggplot(data.sub,aes(x=cut,y=ystar))+
  geom_jitter()+
  facet_wrap(~carat.group)

ggplot(data.sub,aes(x=cut,y=ystar))+
  geom_boxplot()+
  facet_wrap(~carat.group)

```


