---
title: "Guided 4"
author: "Michael Puchalski"
date: "2025-02-11"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
library(tidyverse)
library(palmerpenguins)
df<-penguins
```

### 1. Gentoo Plot of body mass against flipper length

```{r}
gentoo<-df%>%
  filter(species == "Gentoo")

ggplot(gentoo, aes(x=flipper_length_mm, y=body_mass_g))+
  geom_point()+
  labs(x="Flipper Length (mm)", y="Body Mass (g)", title = "Gentoo Body Mass against Flipper Length")

result<-lm(body_mass_g~flipper_length_mm, data = gentoo)
result
```

estimated linear eqaution: y = 54.62(x) - 6787.28

### Question 2

```{r}
confint(result, level = 0.95)[2,]
```

The change in body mass is predicted to be a chnage of 54.62 grams per
mm, The CI where the mean for change per mm will fall between these
values 95% of the time is 44.667 and 64.577.

## Question 3

Hypothesis testing

```{r}
summary(result)
```

Null Hypothesis: H0 = 0; alternate Ha does not equal 0 **p-value** is
very small 2.2 to the -16 power because the p value is so small, the
null should be rejected because it suggest that weight is predicted
rather well by wing length

## Question 4

Yes, in terms of grams in comparison to an animal weighed in kilograms,
the variance of 20 grams that can be found in the confidence interval is
a very miniscule difference.

## Question 5

```{r}
new_flip<- data.frame(flipper_length_mm = 200)
prediction<- predict(result, newdata = new_flip, interval = "confidence", level = 0.95)
prediction
```

Here the predicted weight in grams is 4137.22 with the mean having a 95%
CI of 3954.446 to 4319.993 grams

## Question 6

```{r}
prediction1<- predict(result, newdata = new_flip, interval = "prediction", level = 0.95)
prediction1
```

Lower bound is 3401.121 and upper bound is 4873.319

## Question 7

```{r}
t_stat<- (54.623 - 50)/5.028

t_stat

p_value<- pt(.919, df = 121)
p_value
```

Null Hypothesis: B1 = 50 Alternative Hypothesis: B1 \> 50 test_stat =
0.919 p-value = 0.1798 The p-value is greater than 0.05 meaning we don't
the null hypothesis in favor of our alternative hypothesis
